---
title: "Assignment4"
author: "Rohith chandra koyyala"
date: "20/03/2022"
output:
  html_document:
    df_print: paged
---
loading data and installing libraries
```{r}
Pharmadataset<- read.csv("C:/Users/kramr/Desktop/Pharmaceuticals.csv", header = TRUE)
str(Pharmadataset)
library(tidyverse)
library(factoextra)
library(cluster)
library(ggplot2)
library(gridExtra)
```
removing the missing value which might be present in the data set
```{r}
Pharmadataset <- na.omit(Pharmadataset)
```
accumelating all numerical variables from column 1 to column 9 so that we can cluster all 21 companies
```{r}
row.names(Pharmadataset)<- Pharmadataset[,1]
P<- Pharmadataset[, 3:11]
head(P)
```
scaling the given data using scale function
```{r}
dataframe<- scale(P)
head(dataframe)
```
implementing K means clustering in R for differenet center points
we are going to use multiple values of K and examine the differences in the results
```{r}
kmeans <- kmeans(dataframe, centers = 2, nstart = 30)
kmeans1<- kmeans(dataframe, centers = 5, nstart = 30)
kmeans2<- kmeans(dataframe, centers = 6, nstart = 30)
Plot1<-fviz_cluster(kmeans, data = dataframe)+ggtitle("k=2")
plot2<-fviz_cluster(kmeans1, data = dataframe)+ggtitle("k=5")
plot3<-fviz_cluster(kmeans2, data = dataframe)+ggtitle("k=6")
grid.arrange(Plot1,plot2,plot3, nrow = 2)
```
trying to determine maximum clustering using elbow method
```{r}
dist1<- dist(dataframe, method = "euclidean")# for calculating distance matrix between rows of a data matrix.
fviz_dist(dist1)# Visualizing a distance matrix
```

foe every k, calculate the total within cluster sum of square (wss)
competing and plotting within cluster sum of sqaures fpr k=1 tp k=10
extract wss for 2-25 clusters
the location of the knee in the plot is generally considered as an indicator of the approoriate number of the cluster k= 5
```{r}
set.seed(123)
wss<- function(k){
  kmeans(dataframe, k, nstart =10)$tot.withinss
}
k.values<- 1:10 
wss_clusters<- map_dbl(k.values, wss)
plot(k.values, wss_clusters,
 type="b", pch = 16, frame = TRUE, 
xlab="Number of clusters",
ylab="Total within-clusters sum of squares")
```
final analysis to extract results using 5 cluters and visuals
```{r}
set.seed(123)
final<- kmeans(dataframe, 5, nstart = 25)
print(final)
fviz_cluster(final, data = dataframe)
P%>%
  mutate(Cluster = final$cluster) %>%
  group_by(Cluster)%>% summarise_all("mean")
clusplot(dataframe,final$cluster, color = TRUE, labels = 2,lines = 0)
```
b. interpreting clusters with respect to numerical variable used in clustering
Cluster1 - AHM,SGP,WYE,BMY,AZN, ABT, NVS, LLY
Cluster2 - BAY, CHTT, IVX
Cluster3 - AGN, PHA
Cluster4 - JNJ, MRK, PFE,GSK
Cluster5 - WPI, MRX,ELN,AVE
```{r}
ClusterFormed<- Pharmadataset[,c(12,13,14)]%>% mutate(clusters = final$cluster)%>% arrange(clusters, ascending = TRUE)
ClusterFormed
```
c) is there a pattern in the cluster with respect to numerical variable (10 to 12)?
```{r, message=FALSE, warning=FALSE, fig.width=10}
p1<-ggplot(ClusterFormed, mapping = aes(factor(clusters), fill=Median_Recommendation))+geom_bar(position = 'dodge')+labs(x ='Number of clusters')
p2<- ggplot(ClusterFormed, mapping = aes(factor(clusters),fill = Location))+geom_bar(position = 'dodge')+labs(x ='Number of clusters')
p3<- ggplot(ClusterFormed, mapping = aes(factor(clusters),fill = Exchange))+geom_bar(position = 'dodge')+labs(x ='Number of clusters')
grid.arrange(p1,p2,p3) 
```
As per graph shown above, Cluster1 Suggests to Hold to Moderate Sell
Cluster2 Suggests to Hold 
Cluster3 Suggests to Hold to Moderate Buy
Cluster4 suggests to Hold to Moderate Buy
Cluster5 suggests to Moderate Buy to Moderate Sell

d) giving names to each cluster: 
Cluster1- Sell Cluster,
Cluster2- Hold Cluster,
Cluster3- Buy Cluster,
Cluster4- High Buy Cluster,
Cluster5- Buy Sell Cluster,